Automated reading of residual plots with computer vision models

### 2024-01-29 paper/paper.tex
Auto stash before merge of "master" and "origin/master"

@@

In computer vision models, the input plays a role similar to predictors
in a linear regression model. The quality and relevance of the input
data greatly influence the model's capacity to generate insightful and
meaningful results. A straightforward approach involves feeding the
model a vector of residuals along with a vector of fitted values,
essentially providing all the necessary details for creating a residuals
vs fitted values plot. However, a drawback of this method is the dynamic
input size, as different fitted regression models may have varying
numbers of observations. In modern computer vision models created with
mainstream software like TensorFlow (ref here), the input is typically a
fixed-shape tensor. One solution is to pad the input vector with leading
or trailing zeros if the input tensor expects a longer vector. However,
this approach may fail if the input vector surpasses the designed
length. Another strategy is to summarize the residuals and fitted values
separately using histograms and utilize the counts as the input. By
controlling the number of bins in the histograms, it becomes possible to
provide a fixed-length input vector.

The primary advantage of using an image as input, as opposed to a vector
input format, relates to the development of sophisticated image
processing architectures over the years (see ref here). These
architectures effectively capture and summarize spatial information from
nearby pixels, which is less straightforward with vector input. The only
considerations are the image resolution and the aesthetics of the
residual plot. In general, higher resolution provides more information
to the model but comes with the trade-off of increased complexity and
greater difficulty in training. As for the aesthetics of the residual
plot, a practical solution is to consistently present residual plots in
the same style to the model. This implies that the model can not accept
arbitrary images as input but requires the use of the same preprocessing
software to convert provided residuals and fitted values into a
standardized style for the residual plot.

Other possible input formats include, but are not limited to, a pair of
residual plots, a triplet, and a lineup. Existing computer vision models
designed for image comparison can assess whether a pair of images are
similar or dissimilar (as seen in LeCun's 2005 paper). Applied to our
specific problem, we can define null plots of a fitted model to be
similar to each other, while considering actual residual plots to be
distinct from null plots of any fitted model. A triplet constitutes a
set of three images, denoted as \(image_1\), \(image_2\) and
\(image_3\). In computer vision models, triplets are often used to
predict whether \(image_2\) or \(image_3\) is more similar to
\(image_1\), proving particularly useful for establishing rankings
between samples. In the context of our problem, we can apply the same
criteria to define similarity between images. However, it's important to
note that these two approaches usually require additional considerations
regarding the loss function and, at times, non-standard training
processes due to shared weights between different convolutional blocks.

Presenting a lineup to a model and asking it to predict which residual
plot is the most different one aligns closely with the lineup protocol.
However, if the lineup consists of around 20 plots, similar to previous
human subject experiments (reference here), the resolution of the input
image could become quite large, posing challenges in training the model.
In fact, we experimented with this approach in a pilot study, and the
model's performance was sub-optimal.

We did not explore all the mentioned input formats due to the
considerable costs associated with data preparation and model training.
Considering the implementation cost and the interpretability of the
model, we settled on the single residual plot input format.

@@

Since the Kullback-Leibler divergence can be viewed as the expectation
of the log-likelihood ratio between distribution \(P\) and distribution
\(Q\) evaluated on distribution \(P\), we can reuse the simulated
residuals in matrix \(\boldsymbol{B}\) to estimate the expectation by
the sample mean. With the independence assumption, for non-normal \(P\),
\(KL\) can be estimated by

\begin{align}
\label{eq:kl-3}
\hat{KL} &= \sum_{i = 1}^{n} \hat{KL_i}, \\
\hat{KL_i} &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
\end{align}

\noindent where \(\hat{KL_i}\) is the estimator of the Kullback-Leibler
divergence for an individual residual \(e_i\), \(B_{ij}\) is the
\(i\)-th row and \(j\)-th column entry of the matrix \(B\),
\(\hat{p_i}(.)\) is the kernel density estimator of \(p_i(.)\), \(q(.)\)
is the normal density function with mean zero and an assumed variance
estimated as
\(\hat{\sigma^2} = \sum_{b \in vec(B)}(b - \sum_{b \in vec(B)} b/nm)^2/(nm - 1)\),
and \(vec(.)\) is the vectorization operator which turns a
\(n \times m\) matrix into a \(nm \times 1\) column vector by stacking
the columns of the matrix on top of each other.

@@

\begin{equation}
\label{eq:d-approx}
\hat{D} = cv(plot_{h\times w}(\boldsymbol{e}, \hat{\boldsymbol{y}})),
\end{equation}

\noindent where \(\boldsymbol{e}\) is a vector of residuals obtained
from the fitted model rather than a vector of random variables,
\(\hat{\boldsymbol{y}}\) is the fitted values also obtained from the
fitted model, \(plot(.)_{h \times w}\) is a plotting function that
generates a residuals vs fitted values plot with fixed aesthetic and
then saves it as an image with \(h \times w\) pixels and three colour
channels, and \(cv(.)\) is a computer vision model which takes an
\(h \times w\) image as input and predicts the distance in the domain
\([0, +\infty)\).

@@

Having a computer vision model \(cv\) which gives an approximated
distance between \(0\) and \(+\infty\) is not enough to determine if the
null hypothesis that the model is correctly specified should be
rejected. Theoretically, the distance \(D\) for a correctly specified
model is \(0\), because \(P\) will be the same as \(Q\), and
\(log(1 + 0) = 0\). The computer vision model may not necessary predict
\(0\) for a null plot. Using \ref{fig:false-finding} as an example, it
contains a visual pattern which is an indication of heteroskedasticity.
We would not expect the model to be able to magically tell if the
suspicious pattern is caused by the skewed distribution of the fitted
values or the existence of heteroskedasticity. Assume we only train the
model with this single image, while \(50\)\% of the time the label is
\(0\), and \(50\)\% of the time the label is some constant greater than
\(0\). Then, a perfect model will predict a value between \(0\) and that
constant to achieve a minimum loss. So, it is expected that our computer
vision models will also predict value greater than \(0\) for some null
plots. It will make more sense if we interpret the model prediction as
the visual signal strength of the residual plot. Some null plots could
have outliers or strong visual patterns due to randomness, and the model
will try to summarise those information into the prediction.

To align with the principle of visual inference, we need to provide a
valid null distribution for the test statistic. If we treat the
approximated distance \(\hat{D}\) as a test statistic, then the null
distribution of this statistic can be approximated by the empirical
distribution of \(\hat{D}\) for a large amount of null plots of given a
fitted model. The procedure involves applying the residual rotation
technique \citep{buja2009statistical} on the fitted model to obtain null
residuals. The null residuals are then used to make null plots and fed
into the model. The results are used to construct an empirical
distribution.

@@

With the null distribution of the test statistic,

\begin{itemize}
\tightlist
\item
  How many null samples are needed to get stable 95\% quantile
\item
  How do we make decision based on the quantile (reject or not reject)
\item
  p-value calculation (percentage of null vss \textgreater{} vss)
\end{itemize}

@@

We have mentioned in Section \ref{introduction} that scagnostics
summarise some useful characteristics of the scatter plot. Although the
computer vision model will learn the feature extraction function during
the training process, it will be beneficial to utilize scagnostics to
help computer vision model making more accurate predictions as these
measure could provide additional useful information.

For each generated residual plot, four scagnostics, namely,
``Monotonic'', ``Sparse'', ``Splines'' and ``Striped'' are computed
using the \texttt{cassowaryr} R package (ref here). The computed
measures along with the number of observations of the fitted model are
provided as the second input of the computer vision model.

Other scagnostics are also informative, but unfortunately, they are
unavailable due to a fatal bug in the compiled C program of the
\texttt{interp} R package (ref here) that will crash the process in a
unpredictable manner. For reproducibility, we do not include these
scagnostics in the training data.

@@

\begin{itemize}
\tightlist
\item
  optimization (MSE loss function, with equation)
\item
  a diagram
\item
  explain the existence of removable layers
\item
  explain three different resolutions
\end{itemize}

\hypertarget{training-process-and-hyperparameter-tuning}{%
\subsection{Training process and hyperparameter
tuning}\label{training-process-and-hyperparameter-tuning}}

\begin{itemize}
\tightlist
\item
  software environment (tf and keras)
\item
  tuner (keras\_tuner)
\item
  80\% validation data, 20\% training data
\item
  100 trials (Bayesian optimizer)
\item
  monitor RMSE
\item
  early stopping (50 epochs)
\item
  learning rate reduction (10 epochs, 0.5 factor)
\item
  max 2000 epochs (no one reach)
\item
  restore best epoch of the best trial
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  RMSE for the test set
\item
  R\^{}2
\item
  Mean bias deviation to understand the overall bias
\item
  quantile loss to understand how well the model captures the entire
  distribution of D
\item
  Percentage of predictions within a tolerance interval (like 0.1)
\end{itemize}

@@

\hypertarget{best-model-performance}{%
\subsection{Best model performance}\label{best-model-performance}}

\begin{itemize}
\tightlist
\item
  Metrics for model performance
\item
  Shap values
\item
  Heatmap
\end{itemize}

@@

@@

\begin{itemize}
\tightlist
\item
  power comparison
\item
  decisions
\end{itemize}

>>>>

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

In computer vision models, the quality and relevance of the input data
greatly influence the model's capacity to generate insightful and
meaningful results. The input of the model can be designed in different
formats.

A straightforward approach involves feeding the model a vector of
residuals along with a vector of fitted values, essentially providing
all the necessary information for creating a residuals vs fitted values
plot. However, a drawback of this method is the dynamic input size, as
different fitted regression models may have varying numbers of
observations. For modern computer vision models implemented with
mainstream software like TensorFlow \citep{abadi2016tensorflow}, the
input shape is typically fixed. One solution is to pad the input vectors
with leading or trailing zeros if the input tensor expects longer
vectors, but it may fail if the input vector surpasses the designed
length. Another strategy is to summarize the residuals and fitted values
separately using histograms and utilize the counts as the input. By
controlling the number of bins in the histograms, it becomes possible to
provide fixed-length input vectors.

Another approach is to use an image as input. The primary advantage of
using an image, as opposed to the vector format, is the utilization of
the sophisticated image processing architectures developed over the
years, such as the VGG16 architecture proposed in
\citet{simonyan2014very}. These architectures can effectively capture
and summarize spatial information from nearby pixels, which is less
straightforward with vector input. The only considerations are the image
resolution and the aesthetics of the residual plot. In general, higher
resolution provides more information to the model but comes with the
trade-off of increased complexity and greater difficulty in training. As
for the aesthetics of the residual plot, a practical solution is to
consistently present residual plots in the same style to the model. This
implies that the model can not accept arbitrary images as input but
requires the use of the same preprocessing software to convert provided
residuals and fitted values into a standardized style for the residual
plot.

Other possible input formats include, but are not limited to, a pair of
residual plots, a triplet, and a lineup. \citet{chopra2005learning} has
shown that computer vision models designed for image comparison can
assess whether a pair of images are similar or dissimilar. Applied to
our specific problem, we can define null plots of a fitted model to be
similar to each other, while considering actual residual plots to be
distinct from null plots of any fitted model. A triplet constitutes a
set of three images, denoted as \(image_1\), \(image_2\) and
\(image_3\). It is often used to predict whether \(image_2\) or
\(image_3\) is more similar to \(image_1\), proving particularly useful
for establishing rankings between samples. However, it's important to
note that these two approaches usually require additional considerations
regarding the loss function and, at times, non-standard training
processes due to shared weights between different convolutional blocks.

Presenting a lineup to a model and asking it to predict which residual
plot is the most different one aligns closely with the lineup protocol.
However, if a lineup consists of a large amount of residual plots, the
resolution of the input image could become quite large, posing
challenges in training the model. This approach was experimented in a
pilot study conducted by us but the performance of the trained model was
sub-optimal.

Not all the mentioned input formats were explored and tested in this
study due to the considerable costs associated with data preparation and
model training. The single residual plot input format was chosen for its
relatively low implementation cost and high interpretability.

@@

Since the Kullback-Leibler divergence can be viewed as the expectation
of the log-likelihood ratio between distribution \(P\) and distribution
\(Q\) evaluated on distribution \(P\), we can reuse the simulated
residuals in matrix \(\boldsymbol{B}\) to estimate the expectation by
the sample mean. With the independence assumption, for non-normal \(P\),
\(D_{KL}\) can be estimated by

\begin{align}
\label{eq:kl-3}
\widehat{D_{KL}} &= \sum_{i = 1}^{n} \widehat{D_{KL}^{(i)}}, \\
\widehat{D_{KL}^{(i)}} &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
\end{align}

\noindent where \(\widehat{D_{KL}^{(i)}}\) is the estimator of the
Kullback-Leibler divergence for an individual residual \(e_i\),
\(B_{ij}\) is the \(i\)-th row and \(j\)-th column entry of the matrix
\(B\), \(\hat{p_i}(.)\) is the kernel density estimator of \(p_i(.)\),
\(q(.)\) is the normal density function with mean zero and an assumed
variance estimated as
\(\widehat{\sigma^2} = \sum_{b \in vec(B)}(b - \sum_{b \in vec(B)} b/nm)^2/(nm - 1)\),
and \(vec(.)\) is the vectorization operator which turns a
\(n \times m\) matrix into a \(nm \times 1\) column vector by stacking
the columns of the matrix on top of each other.

@@

\begin{equation}
\label{eq:d-approx}
\hat{D} = f(V(\hat{\boldsymbol{e}}, \hat{\boldsymbol{y}})),
\end{equation}

\noindent where \(\hat{\boldsymbol{e}}\) is a vector of residuals
obtained from the fitted model rather than a vector of random variables,
\(\hat{\boldsymbol{y}}\) is the fitted values also obtained from the
fitted model, \(V(.)\) is a plotting function that generates a residuals
vs fitted values plot with fixed aesthetic and then saves it as an image
with \(h \times w\) pixels and three colour channels, and \(f(.)\) is a
computer vision model which takes an \(h \times w\) image as input and
predicts the distance in the domain \([0, +\infty)\).

@@

Theoretically, the distance \(D\) for a correctly specified model is
\(0\), because \(P\) will be the same as \(Q\). However, the computer
vision model may not necessary predict \(0\) for a null plot. Using
Figure \ref{fig:false-finding} as an example, it contains a visual
pattern which is an indication of heteroskedasticity. We would not
expect the model to be able to magically tell if the suspicious pattern
is caused by the skewed distribution of the fitted values or the
existence of heteroskedasticity. Assume we only train the model with
this single image, while \(50\)\% of the time the label is \(0\), and
\(50\)\% of the time the label is some constant \(c > 0\). Then, a
perfect model will predict \(0 < \hat{D} < c\) to achieve a minimum
loss. Some null plots could have outliers or strong visual patterns due
to randomness, and a reasonable model will try to summarise those
information into the prediction, resulting in \(\hat{D} > 0\).

This property is not an issue if \(\hat{D} \gg 0\) for which the visual
signal of the residual plot is very strong, and we usually do not need
any further examination of the significance of the result. However, if
the visual pattern is weak or moderate, having \(\hat{D}\) will not be
sufficient to determine if the null hypothesis that the model is
correctly specified should be rejected.

To solve this issue while aligning with the principle of visual
inference, the approximated distance \(\hat{D}\) can be defined as a
test statistic. And the null distribution of this statistic can be
approximated by the empirical distribution of \(\widehat{D_{null}}\) for
a large amount of null plots of a fitted model. The procedure involves
applying the residual rotation technique \citep{buja2009statistical} on
the fitted model to obtain null residuals. The null residuals are then
used to make null plots and fed into the model. The predicted distance
are used to construct an empirical distribution.

@@

Let \(\widehat{D_{null}^{(i)}}\) be the predicted distance of the
\(i\)-th null plots, where \(i = 1,...,m\) and \(m \in \mathbb{N}^+\) is
a sufficiently large number. Quantiles of the null distribution can then
be estimated using the sample quantiles available in statistical
software such as R. The details of the sample quantile computation can
be found in \citet{hyndman1996sample}. In statistical testing, analysts
often care about certain quantiles of the null distribution, such as
90\% quantile, 95\% quantile and 99\% quantile. These quantiles are used
as thresholds to decide if \(H_0\) needs to be rejected. For example, if
\(\hat{D}\) is greater than and equal to the 95\% sample quantile of the
null distribution, we could say the predicted distance for the actual
residual plot is significantly different from the predicted distance for
null plots with 95\% significance level. Based on our experience, in
order to get a stable estimate of the 95\% quantile, \(m\) usually needs
to be at least \(100\). And if the null distribution has a long tail,
more null plots will be needed. Alternatively, a p-value can be used to
represents the probability of observing an event equally or more extreme
than the given event under \(H_0\), and it can be estimated by
\(1/m\sum_{i=1}^{m}I\left(\widehat{D_{null}^{(i)}} \geq \hat{D}\right)\).

@@

We have mentioned in Section \ref{introduction} that scagnostics are a
set of manually designed visual feature extraction functions. Although
our computer vision model will learn its own feature extraction function
during the training process, it will be beneficial to utilize additional
information provided by scagnostics to help computer vision model making
more accurate predictions.

For each generated residual plot, four scagnostics, namely,
``Monotonic'', ``Sparse'', ``Splines'' and ``Striped'' are computed
using the \texttt{cassowaryr} R package (ref here). The computed
measures along with the number of observations of the fitted model are
provided as the second input of the computer vision model. Other
scagnositcs are also informative, but unfortunately, they are
unavailable due to a fatal bug in the compiled C program of the
\texttt{interp} R package (ref here) that will crash the process in a
unpredictable manner. For reproducibility, we do not include these
scagnostics in the training data.

@@

The architecture of the computer vision model is adopted from a existing
architecture called VGG16 that proven to be performant in image
classification \citep{simonyan2014very}. The first layer is the input
layer with shape \(n \times h \times w \times 3\) which could accept
\(n\) RGB images, followed by a grayscale conversion layer using the
luma formula under the CCIR 601 standard (ref here) to convert the
colour image to an grey scale image. Grayscale is enough for our task
since data points are plotted in black. Three different combinations of
\(h\) and \(w\) are chosen to seek a sufficient large image resolution
for this problem. This includes \(32 \times 32\), \(64 \times 64\) and
\(128 \times 128\).

The processed image is then used as the input of the first
convoluational block. The model has at most five consecutive
convolutional blocks which is similar to the original VGG16
architecture. In each block, there are two 2D convoluational layers
followed by two activation layers respectively, and there is a 2D max
pooling layer follows the second activation layer. The 2D convolutional
layer convolves the input with a fixed number of \(3 \times 3\)
convolution filters, and the 2D max pooling layer downsamples the input
along its spatial dimensions by taking the maximum value over a
\(2 \times 2\) window for each channel of the input. The activation
layer is set to be the rectified linear unit activation function (ReLu),
which is a standard practice in deep learning. This introduces a
non-linear transformation of the output of the 2D convolutional layer.
Furthermore, to regularize the training, a batch normalization layer is
added after each of the 2D convolutional layer and before the activation
layer. A dropout layer is also appended at the end of each convolutional
block to randomly set some inputs to zeros during training.

The output of the last convolutional block will be summarised by a
global max pooling layer or a global average pooling layer to gain a
two-dimensional tensor. Since we want to utilize the information
contained in scagnostics, the two-dimensional tensor is concatenated
with an additional \(n \times 5\) tensor, which contains the
``Monotonic'', ``Sparse'', ``Splines'' and ``Striped'' measures and the
number of observations for \(n\) residual plots.

The concatenated tensor will be fed into the final prediction block.
This block contains two fully-connected layers, the first one has at
least \(128\) units followed by a dropout layer. Sometimes, a batch
normalization layer will be inserted between the fully-connected layer
and the dropout layer. The second fully-connected layer has only one
unit, serves as the output of the model.

The model weights are randomly initialized and they are optimized by the
Adam optimizer with the mean square error loss function. Therefore, we
try to find a set of weights \(\hat{\boldsymbol{\theta}}\) by

\[\hat{\boldsymbol{\theta}} = \underset{\boldsymbol{\theta}}{argmin}\frac{1}{n_{train}}\sum_{i=1}^{n_{train}}(D_i - f_{\boldsymbol{\theta}}(V_i, S_i))^2,\]

\noindent where \(n_{train}\) is the number of training samples, \(V_i\)
is the \(i\)-th residual plot and \(S_i\) is the additional information
about the \(i\)-th residual plot including four scagnostics and the
number of observations.

\begin{itemize}
\tightlist
\item
  a diagram for the model
\end{itemize}

\hypertarget{training-and-hyperparameter-tuning}{%
\subsection{Training and hyperparameter
tuning}\label{training-and-hyperparameter-tuning}}

To obtain an optimal deep learning model, hyperparameters such as
learning rate often needs to be tuned via a tuner. In this study, we
used the Bayesian optimization tuner from the \texttt{keras-tuner}
Python library (ref here) to tune the hyperparameters. The complete list
of hyperparameters is provided in Table \ref{tab:hyperparameter}.

The number of base filters determines the number of filters for the
first 2D convolutional layer. In VGG16, the number of filters for the 2D
convolutional layer in a block is 2 times the number in the previous
block, except for the last block which has the same number of
convolution filters as the previous one. This hyperparameter helps to
control the complexity of the computer vision model. More base filters
will result in more trainable parameters.

Similarly, the number of units for the first fully-connected layer
determines the complexity of the final prediction block. And more units
will lead to more trainable parameters.

Dropout rate and batch normalization are flexible hyperparameters which
works jointly with other hyperparameters to smooth the training process.
A high dropout rate is needed when the model learns too much noise from
the training dataset. And a low dropout rate is needed when the model is
complex and difficult to converge. Batch normalization on the other hand
may be needed for solving the internal covariate shift problem due to
randomness in the weight initialization and randomness in the input
data.

Having additional inputs including the scagnostics and the number of
observations could be beneficial for making more accurate prediction. So
we let the tuner to decide if these inputs are actually needed.

Learning rate is one of the most important hyperparameters, and it
determines the step size of the optimization algorithm. A high learning
rate could help the model to avoid getting stuck in a local minimum, but
it could also ruin the optimization process by jumping over the global
minimum. A low learning rate could smooth the training process but it
increases the time to converge and the chance of getting stuck in local
minimum.

The model is trained on a high-performance computing platform owned by
Monash university with TensorFlow (ref here) and Keras (ref here).
During the training process, 80\% of the training data are used as the
actual training data, and 20\% of the training data are reserved as
validation data. We let the Bayesian optimization tuner runs 100 trials
to find the best hyperparameter values according to the validation root
mean square error. The tuner will restore the best epoch of the best
model from the trials. Early stopping is also applied such that the
training process will halt if the validation root mean square error does
not improve for 50 epochs. The maximum allowed epochs is 2000 where no
models reach.

\begin{table}

\caption{\label{tab:hyperparameter}Name of hyperparameters and their correspoding domain for the computer vision model.}
\centering
\begin{tabular}[t]{ll}
\toprule
Hyperparameter & Domain\\
\midrule
Number of base filters & \{4, 8, 16, 32, 64\}\\
Dropout rate for convolutional blocks & {}[0.1, 0.6]\\
Batch normalization for convolutional blocks & \{false, true\}\\
Type of global pooling & \{max, average\}\\
Ignore additional inputs & \{false, true\}\\
\addlinespace
Number of units for the fully-connected layer & \{128, 256, 512, 1024, 2048\}\\
Batch normalization for the fully-connected layer & \{false, true\}\\
Dropout rate for the fully-connected layer & {}[0.1, 0.6]\\
Learning rate & {}[1e-8, 1e-1]\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{itemize}
\item
  RMSE for the test set
\item
  R\^{}2
\item
  Mean bias deviation to understand the overall bias
\item
  quantile loss to understand how well the model captures the entire
  distribution of D
\item
  Percentage of predictions within a tolerance interval (like 0.1)
\item
  comparison with human visual inference
\item
  overview of the human subject experiment
\item
  metrics
\item
  agreement rate
\item
  check cases that are aggreed
\item
  check cases that are disaggreed
\end{itemize}

@@

\hypertarget{best-hyperparameters}{%
\subsection{Best hyperparameters}\label{best-hyperparameters}}

Based on the tuning process described in Section
\ref{training-and-hyperparameter-tuning}, the best hyperparameter values
are provided in Table \ref{tab:best-hyperparameter}. It can be observed
that at least \(32\) base filters are needed, and the preferable choice
for the \(64 \times 64\) model and the \(128 \times 128\) model is
\(64\) base filters, which is the same as the original VGG16
architecture. The best values for dropout rate for convolutional blocks
is around \(0.4\). And applying batch normalization for convolutional
blocks has a positive impact on the performance. All best models choose
to preserve the additional inputs to reduce the validation error. The
number of units needed for the fully-connected layer is \(256\), which
is not a very large number comparing to the VGG16 classifier, suggesting
that the problem we are trying to solve is relatively easy. The best
learning rates are usually smaller than the default value recommended by
Keras which is 0.001.

\begin{table}

\caption{\label{tab:best-hyperparameter}Hyperparameters values for the best computer vision models with differnet input sizes.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llll}
\toprule
Hyperparameter & $32 \times 32$ & $64 \times 64$ & $128 \times 128$\\
\midrule
Number of base filters & 32 & 64 & 64\\
Dropout rate for convolutional blocks & 0.4 & 0.3 & 0.4\\
Batch normalization for convolutional blocks & true & true & true\\
Type of global pooling & max & average & average\\
Ignore additional inputs & false & false & false\\
\addlinespace
Number of units for the fully-connected layer & 256 & 256 & 256\\
Batch normalization for the fully-connected layer & false & true & true\\
Dropout rate for the fully-connected layer & 0.2 & 0.4 & 0.1\\
Learning rate & 0.0003 & 0.0006 & 0.0052\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{model-performance}{%
\subsection{Model performance}\label{model-performance}}

The training and test performance of models with three different input
sizes are provided in Table \ref{tab:performance}. Among these models,
the \(64 \times 64\) model and the \(32 \times 32\) model are
consistently having the best metrics on the training set and the test
set respectively. The mean absolute error shows that the difference
between \(\hat{D}\) and \(D\) is around \(0.43\) on the test set, which
is very small considering the range of \(D\) is normally between \(0\)
and \(7\). The high \(R^2\) also suggests the prediction is mostly
linearly correlated with the target.

Figure \ref{fig:model-performance} displays a scatter plot for
\(D - \hat{D}\) vs \(D\). For residual plots with \(D\) close to zero,
all the models tends to over-predict \(\hat{D}\), leading to negative
residuals. When \(D\) is between \(1.5\) and \(6\), the smooth curves
indicate that \(\hat{D}\) is not heavily biased, and there are some
instances where \(\hat{D}\) is severely under-predicted. And when \(D\)
is greater than \(6\), there are more instances being under-predicted
than over-predicted. It can be observed from Figure
\ref{fig:model-performance-density} that the models produced much more
\(\hat{D}\) at around \(1\) than expected.

Given the model performance metrics, only the best model evaluated on
the test set, i.e.~the \(32 \times 32\) model, will be used in the
following analysis.

\begin{table}

\caption{\label{tab:performance}The training and test performance of three models with different input sizes.}
\centering
\begin{tabular}[t]{lllll}
\toprule
 & RMSE & $R^2$ & MAE & Huber loss\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Training set}}\\
\hspace{1em}$32 \times 32$ & 0.531 & 0.937 & 0.364 & 0.126\\
\hspace{1em}$64 \times 64$ & 0.405 & 0.963 & 0.260 & 0.072\\
\hspace{1em}$128 \times 128$ & 0.432 & 0.959 & 0.290 & 0.084\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Test set}}\\
\hspace{1em}$32 \times 32$ & 0.660 & 0.901 & 0.434 & 0.181\\
\hspace{1em}$64 \times 64$ & 0.674 & 0.897 & 0.438 & 0.186\\
\hspace{1em}$128 \times 128$ & 0.692 & 0.892 & 0.460 & 0.199\\
\bottomrule
\end{tabular}
\end{table}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/model-performance-1}

\caption{Residuals vs distance plot on training and test data for three best models with different input sizes. The blue lines are smoothing curves produced by fitting gnealized additive models.}\label{fig:model-performance}
\end{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/model-performance-density-1}

\caption{Density plot of predicted distance on training and test data for three best models with different input sizes. The density drawn in red is the density of the distance.}\label{fig:model-performance-density}
\end{figure}

@@

\begin{itemize}
\tightlist
\item
  basic background
\item
  highlights the same and the different settings
\end{itemize}

@@

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-2-1}

\caption{Residuals vs distance plot on data used in the human subject experiment. The blue line is a smoothing curve produced by fitting a gnealized additive model.}\label{fig:unnamed-chunk-2}
\end{figure}

\begin{table}

\caption{\label{tab:experiment-performance}The performance of the $32 \times 32$ model on the data used in the human subject experiment.}
\centering
\begin{tabular}[t]{lllll}
\toprule
Type & RMSE & $R^2$ & MAE & Huber loss\\
\midrule
heteroskedasticity & 0.768 & 0.863 & 0.594 & 0.263\\
non-linearity & 0.731 & 0.817 & 0.560 & 0.242\\
\bottomrule
\end{tabular}
\end{table}

It can be observed from Table \ref{tab:experiment-performance} that all
the performance metrics are lower than the metrics evaluated on the test
data.
